A template for the \emph{arxiv} style

### 2021-09-01 paper-rmd/paper.tex
added rmd files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Untitled probability judgment paper draft},
  pdfauthor={Derek Powell1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{DRAFT}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Untitled probability judgment paper draft}
\author{Derek Powell\textsuperscript{1}}
\date{}

\authornote{

Arizona State University, School of Social and Behavioral Sciences

The authors made the following contributions. Derek Powell: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Arizona State University, School of Social and Behavioral Sciences}

\abstract{
Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic. However, people's explicit probability judgments often incoherently violate the laws of probability. Two recent proposals, the ``Probability Theory plus Noise'' (Costello \& Watts, 2014) and ``Bayesian Sampler'' (Zhu, Sanborn, \& Chater, 2020) theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic. The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity. Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the \(LOO_{ic}\) information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020). My findings support the Bayesian Sampler's account of conditional probability judgments, and I find the data are best explained by a ``hybrid'' model that can be seen as a modification of both theories.
}

@@

Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) with remarkable successes across domains, including perception (e.g. Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012). At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule. It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''

Trouble starts as soon as we seek to measure beliefs. According to Bayesian theories of cognition and epistemology (Jaynes, 2003), the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities. So asking people to express beliefs in terms of probability seems only natural.

Unfortunately, people's explicit probability judgments routinely violate the axioms of probability theory. For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory (Tversky \& Kahneman, 1983). Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other effects illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013). Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.

Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020). Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).

@@

where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}). The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics (Costello \& Watts, 2014, 2016, 2017, 2018).

Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model of probability judgment they call the ``Bayesian Sampler.'' Under this model, probability judgment is itself seen as a process of Bayesian inference. To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event. Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment. The effect of this prior is to regularize the judgments, pushing them away from extreme values. They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.~encourage taking large risks based on a small number of mental samples. However, this hedging comes at the cost of systematic incoherence and biases. Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).

@@

By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020). However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project. Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence. So, the conditioning of beliefs is at the heart of these theories.

According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments. By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models (Chater et al., 2020).

In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian (Costello \& Watts, 2016). According to the PT+N model, conditional probabilities P(A\textbar B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B. The PT+N model predicts conditional probability estimates using the following equation:

\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]

The PT+N's account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally. In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.

@@

Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments. Unfortunately, their results were somewhat equivocal. They fit both ``simple'' and ``complex'' versions of each model and computed a Bayesian Information Criteria score (BIC) for all models. The complex versions of these models introduce additional parameters d' and N' that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments. These additional parameters are needed in order to explain conjunction fallacies (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020) . Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).

@@

In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better). It is not obvious what conclusions should be drawn from these results. Of course, the simple Bayesian Sampler model appears to win by the numbers. But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020). So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory. For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together. This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.

@@

When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to ``overfit'' the data. Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data (Gelman, Hwang, \& Vehtari, 2014). Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models' complexity.

There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models. First, the models differ not only in the number of parameters but in the domain of those parameters. Zhu and colleagues (2020) restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}. This restriction constrains the Bayesian Sampler model's prior distribution to the class of ``ignorance priors.'' Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]} whereas the PT+N model permits noise values between {[}0, 0.50{]}.

Second, it is not immediately clear how the models' structural differences impact their flexibility. To be sure, Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities. But, what impact does the PT+N model's treatment of conditional probabilities have on model complexity? Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized? Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility? Determining model complexity a priori is not always straightforward when the models being compared differ structurally.

Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant. Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants. In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term. Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.

@@

Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison. A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\). These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data. Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017). In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling. This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.

@@

Experiment 1 asked about the events {[}icy, frosty{]} and {[}normal, typical{]} (e.g.~``what is the probability that the weather in London is normal and not typical?''). The authors' goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated. Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment. This is especially clear, as the authors note, for disjunctive trials such as ``normal or typical,'' where ``or typical'' might not be read as a disjunction but rather an elaborative clause. In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.

Experiment 2 focused on more moderately correlated events, {[}cold, rainy{]} and {[}windy, cloudy{]}, that do not admit these misinterpretations. In addition, a third experimental condition asking about {[}warm, snowy{]} was also included in the experiment, but was dropped from the analyses reported in the paper. Exploring the raw responses from this condition reveals a substantial fraction of ``zero'' and ``one'' responses for certain trials. This may reflect a different response process than was intended. For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero---failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day. Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.

@@

Notice that for each model the probability judgments depend on underlying subjective probabilities, derived from a mental sampling process. These subjective probabilities are unobserved, and must be estimated as a latent variable. Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (\(A \land B\), \(\neg A \land B\), \(A \land \neg B\), \(\neg A \land \neg B\)).

@@

\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
\[\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) \]
\[ d_j = \frac{expit(\delta_j)}{2}\]
\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
\[ \delta_j \sim Normal(0,1)\]
\[ \Delta\delta_j \sim Normal(0,1)\]
\[ k \sim HalfCauchy(20)\]

@@

Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model. To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its \(\beta\), N, and N' parameters. I constrain d to {[}0, \(1/3\){]} for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\). This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.

@@

>>>>

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{A template for the \emph{arxiv} style}

\author{
    Derek Powell
   \\
    School of Social and Behavioral Sciences \\
    Arizona State University \\
  Phoenix, AZ \\
  \texttt{\href{mailto:dmpowell@asu.edu}{\nolinkurl{dmpowell@asu.edu}}} \\
  }

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\def\tightlist{}

\begin{abstract}
Enter the text of your abstract here.
\end{abstract}

\keywords{
    blah
   \and
    blee
   \and
    bloo
   \and
    these are optional and can be removed
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Bayesian theories of cognition have had remarkable successes in
explaining human reasoning and behavior across many domains (big cite).
{[}The core of these theories is that people reason according to
subjective mentally-represented degrees of belief, and they specify how
they should be revised in light of evidence. It is somewhat embarrassing
then that one area where these theories seem to fall down is in
describing human ``beliefs'\,' of the simple and everyday sort, such as
beliefs like''it will rain tomorrow``, ``vaccines are safe,'' or ``this
politician is trustworthy.''

Trouble starts as soon as we seek to measure beliefs. According to
Bayesian theories of cognition and epistemology, the degree to which
people believe in various propositions should reflect subjective mental
probabilities. So asking people to express beliefs in terms of
probability seems only natural.

Unfortunately, people's explicit probability judgments routinely violate
the axioms of probability theory. For example, human probability
judgments often exhibit the ``conjunction fallacy'': people will often
judge the conjunction of two events (e.g.~``Tom Brady likes football and
miniature horses'') as being more probable than one of the events in
isolation (e.g.~``Tom Brady likes miniature horses''), a plain and
flagrant violation of probability theory (cite some examples). Other
demonstrations of the incoherence of probability judgments include
disjunction fallacies (e.g.~XXX), ``unpacking'' effects (e.g.~fox \&
tversky), and a variety of other effects illustrating the incoherence of
human probability judgments {[}cite{]}. Altogether these findings have
led many researchers to abandon the notion that credences are
represented as probabilities.

Recently however, two groups of researchers have proposed theories of
human probability judgments that account for biases and apparent
incoherence in these judgments while maintaining that mental credences
are fundamentally probabilistic (Costello and Watts 2014; Zhu, Sanborn,
and Chater 2020). Both of these theories build on the increasingly
popular notion that a variety of human reasoning tasks are accomplished
by drawing a limited number of samples from probabilistic mental models
(see also Chater et al. 2020; Dasgupta, Schulz, and Gershman 2017).

@@

where \emph{d} represents the probability with which samples will be
misread, or the amount of noise in the judgment process (where by
assumption a maximum of 50\% of samples can be misread, so that d is a
number in the range {[}0, .50{]}). The PT+N theory provides a unified
accounts for a wide variety of biases in probability judgment that were
previously attributed to different types of heuristics (Costello and
Watts 2014, 2016, 2017, 2018).

Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model
of probability judgment they call the ``Bayesian Sampler.'' Under this
model, probability judgment is itself seen as a process of Bayesian
inference. To judge the probability of an event, a limited number of
samples are again drawn from a mental model of the event. Then, those
``observed'' samples are integrated with a prior over probabilities to
produce a probability judgment. The effect of this prior is to
regularize the judgments, pushing them away from extreme values. They
argue this regularization of judgments is adaptive, in that it pushes
judgments and decisions away from extremes that might, e.g.~encourage
taking large risks based on a small number of mental samples. However,
this hedging comes at the cost of systematic incoherence and biases.
Like the PT+N model, their proposed model accounts for a wide array of
biases in probability judgments, including the novel biases identified
by Costello and Watts (Costello and Watts 2014, 2016).

@@

By explaining the incoherence of human probability judgments using
coherent mental probabilities, both models have the potential to rescue
the larger project of Bayesian cognitive science as applied to everyday
beliefs (Chater et al. 2020). However, because they diverge in their
treatment of conditional probability judgments, the two theories fit
very differently into the larger Bayesian project. According to the
Bayesian sampler model, conditioning is something that happens in the
mental model of the events, not as part of the process of rendering
probability judgments. Thus, the Bayesian sampler model does not assign
any special status to conditional probability judgments and so fits
neatly into the larger Bayesian project of cognitive science:
probability judgments are simply another judgment process applied to the
outputs of other Bayesian mental models (Chater et al. 2020).

In contrast, the PT+N model presents a constructive account of
conditional probability judgments that is fundamentally non-Bayesian
(Costello and Watts 2016). According to the PT+N model, conditional
probabilities P(A\textbar B) are estimated by a two-stage sampling
procedure: first both events A and B are sampled with noise, and then a
second noisy process computes the ratio of the events read as A and B
over events read as B. The PT+N model predicts conditional probability
estimates using the following equation:

\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]\\
Given that conditional inference and belief updating lies at the core of
Bayesian cognitive theory, PT+N's account of conditional probability
judgment separates it from the Bayesian sampler theory not just
quantitatively, but quite fundamentally. First, it denies any central
importance of conditional probability judgments by pulling conditioning
out of the mental model of an event and placing it into the probability
judgment process. And perhaps more importantly, this non-bayesian
account of conditional inference can no longer be considered
``rational'' at any stage under the PT+N model.

@@

Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model
against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as
explanations for human probability judgments in two experiments.
Unfortunately, their results were somewhat equivocal. They fit both
``simple'' and ``complex'' versions of each model and computed a
Bayesian Information Criteria score (BIC) for all models. The complex
versions of these models introduce additional parameters d' and N' that
allow for different patterns of judgments for conjunctive and
disjunctive judgments as compared with simple probability judgments.
These additional parameters are needed in order to explain conjunction
fallacies (Costello \& Watts, 2017; Zhu et al., 2020). Table 1 below
presents the total BIC scores computed for each model as originally fit,
using the authors' original code and saved model outputs (Zhu et al,
2020, Supplementary materials).

@@

In both experiments, the simple version of the Bayesian Sampler scores
best, but the complex version of the PT+N model comes in second (lower
BIC scores are better). It is not obvious what conclusions should be
drawn from these results. Of course, the simple Bayesian Sampler model
appears to win by the numbers. But conjunction fallacies are an
extremely robust empirical finding (e.g.~cite) and are clearly present
in the data collected by Zhu and colleagues (2020). So we might rule out
the simple variants of the models on the grounds that they will fail to
capture important qualitative features of the data, which would instead
favor the PT+N theory. For their part, Zhu and colleagues (2020) elect
to split the baby by averaging the simple and complex model scores
together. This approach somewhat favors the Bayesian Sampler theory
overall, though they are rightfully cautious about drawing any strong
conclusions.

@@

Zhu and colleagues (2020) warn that BIC, which penalizes models based
only on the number of parameters in the model, cannot fully account for
the differences in the models' complexity. There are at least three
challenges to accounting for model complexity in the comparison of these
models. First, the models differ not only in the number of parameters
but in the domain of those parameters. Zhu and colleagues (2020)
restrict the range of the \(\beta\) parameter in their Bayesian Sampler
model to fall within the interval {[}0,1{]}. This restriction constrains
the Bayesian Sampler model's prior distribution to the class of
``ignorance priors,'' symmetric beta distributions where \(\beta < 1\).
Via the bridging conditions (eq X), they show that this restricts the
``noise level'' for the Bayesian Sampler, represented in implied d under
the PT+N model, to {[}0, \(\frac{1}{3}\){]} whereas the PT+N model
permits noise values between {[}0, 0.50{]}.

Second, it is not immediately clear how the structural differences
between the models impact their flexibility. For the purposes of model
comparison, the ``complexity penalty'' that models truly ought to pay is
determined by their relative flexibility in accommodating different
patterns of data, not simply the number or domain of their parameters
(cite Gelman). To be sure, Zhu and colleagues' (2020) bridging condition
makes it clear that the PT+N model is more flexible when it comes to
predicting unconditional probabilities. But, what impact does the PT+N
model's treatment of conditional probabilities have on model complexity?
Is this component of the model a sort of Ptolemaic epicycle, adding
complexity to the model that should be penalized? Or, does it constitute
a commitment to novel predictions that thereby constrain its
flexibility? Determining model complexity a priori is not always
straightforward when the models being compared differ structurally.

Finally, one potential explanation for the relative weakness of the
complex model versions is Zhu and colleagues' (2020) use of unpooled
models, where parameters are estimated independently for each individual
participant. Adding a parameter for each participant may effectively
over-penalize the ``complex'' variants relative to the simple variants,
especially if there is actually limited heterogeneity across
participants. In contrast, a comparison of fully pooled variants of the
simple and complex models (where a single population parameter is used
for all participants) would require adding only one extra parameter to
the penalty term. Partial pooling is a solution that balances between
these extreme approaches, allowing for an accounting of heterogeneity
without over-penalizing in cases where heterogeneity is low.

@@

Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian
data analysis framework that permits a more decisive (or at least
alternative) comparison. A Bayesian analysis allows issues of model
complexity to be addressed through comparisons of model fit based on
information criteria, such as WAIC and LOO\textsubscript{ic}. These
indices allow for complexity penalties that directly capture differences
in the flexibility of the models, rather than being estimated based
solely on the number of distinct parameters {[}cite gelman{]}. In
addition, this framework also supports straightforward implementation of
hierarchical versions of these models with partial pooling. This allows
for information about model parameters to be shared across participants,
resulting in an overall reduction in model complexity and a more
realistic test of the models.

@@

Experiment 1 asked about the events \{icy, frosty\} and \{normal,
typical\} (e.g.~``what is the probability that the weather in London is
normal and not typical?''). The authors' goal was to ask about highly
correlated events, but the events used are perhaps nearly perfectly
correlated. Because the terms used to describe these events are nearly
synonymous, there is a concern about the interpretation of the
statements evaluated in this experiment. This is especially clear, as
the authors note, for disjunctive trials such as ``normal or typical,''
where ``or typical'' might not be read as a disjunction but rather an
elaborative clause. In light of these concerns, I excluded the
disjunctive trials from Experiment 1 from my analyses. Experiment 2
focused on more moderately correlated events, \{cold, rainy\} and
\{windy, cloudy\}, that do not admit these misinterpretations. In
addition, a third experimental condition asking about \{warm, snowy\}
was also included in the experiment, but was dropped from the analyses
reported in the paper. Exploring the raw responses from this condition
reveals a substantial fraction of ``zero'' and ``one'' responses for
certain trials. This may reflect a different response process than was
intended. For instance, some participants may have engaged in deductive
reasoning to judge that it is not possible for the weather to at once be
warm and snowy, and therefore responded with zero---failing to properly
consider that it is possible (at least logically) for it to be warm and
snowy at different times within the same day. Given these potentially
aberrant responses, I followed Zhu and colleagues (2020) in ignoring
data from this condition.

@@

Notice that for each model the probability judgments depend on
underlying subjective probabilities, derived from a mental sampling
process. These subjective probabilities are unobserved, and must be
estimated as a latent variable. Here, they are estimated as a
four-dimensional dirichlet distribution for each subject, representing
the probability of the elementary events (A\&B, \textasciitilde A\&B,
A\&\textasciitilde B, \textasciitilde A\&\textasciitilde B).

@@

\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
\[\mu_{ij} = f_{PT+N}(p_i, d_j, d_j) \]
\[ d_j = \frac{expit(\delta_j)}{2}\]
\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
\[ p_i \sim Dirichlet(\overrightarrow{1})\]
\[ \delta_j \sim Normal(0,1)\] \[ \Delta\delta_j \sim Normal(0,1)\]
\[ k \sim HalfCauchy(20)\]

@@

Recall that Zhu and colleagues (2020) identified a bridging condition
relating b and N in the Bayesian Sampler model to the d parameter of the
PT+N model. To support direct comparisons of the models, I parameterize
the Bayesian Sampler model according to the implied d and d', rather
than directly according to its \(\beta\), N, and N' parameters. I
constrain d to {[}0, \(\frac{1}{3}\){]} for the Bayesian Sampler model
to reflect the assumption that \(\beta \in [0, 1]\). This allows the
same priors to be used for the corresponding Bayesian Sampler and PT+N
models, simplifying their comparison.

@@

\bibliographystyle{unsrt}
\bibliography{references.bib}
